### 2021-11-12 caseForSupport.tex
Splitted caseForSupport.tex into several files

\documentclass[a4paper,11point]{article}
\usepackage[margin=2cm]{geometry}

\usepackage[colorlinks=true]{hyperref}
\usepackage{natbib}
\usepackage{apalike}
\usepackage{graphicx}
\usepackage{verbatim}

\makeatletter
\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
            {-2.5ex\@plus -1ex \@minus -.25ex}%
            {1.25ex \@plus .25ex}%
            {\normalfont\normalsize\bfseries}}
\makeatother
\setcounter{secnumdepth}{4} % how many sectioning levels to assign numbers to
\setcounter{tocdepth}{4}    % how many sectioning levels to show in ToC

\title{Case for Support}
\author{}

\begin{document}

\maketitle

\tableofcontents

\newpage

\setcounter{page}{1}

\section{Background to the resource}

\input{backgroundToTheResource}

\section{Details of resource}

\input{detailsOfResource}

\section{Community demand}
\label{sec:communityDemand}

\input{communityDemand}

\section{User engagement}
\label{sec:userEngagement}

\input{userEngagement}

\section{Resource management}

\section{Long term sustainability planning}

\section{Potential for economic and societal impact}

\section{Track Record}

\bibliographystyle{apalike}
\bibliography{bonsai,epsrcSoftwareForResearchCommunities,monitoringBehavior,stateSpaceModels,machineLearning,gaussianProcesses}

\end{document}