### 2021-11-12 bonsai.tex

Bonsai~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
programming language that emphasizes performance, flexibility, and ease-of-use,
allowing scientists with no previous programming experience to quickly develop
their own high-performance data acquisition and experimental control systems
(Figure 1c). Bonsai combines a high-level event algebra for data streams with
an integrated development environment (IDE) and a strong library of plugins
supporting multiple hardware and software packages used by the biomedical
research community (Figures 1a, 1b).

Standard software tools for data analysis (ImageJ, MATLAB, R, etc.) have been
transformative to the progress of increasingly “data-rich” sciences. However,
the equivalent standardized software tools for data acquisition and control of
animal behaviour experiments are still lacking. Life science experiments demand
a combination of multiple instrumentation and control technologies, in both
physiology and behaviour studies. The growing complexity on both the amount of
data that is collected, and the rich conditions under which behaviour must be
explored, place an increasing burden on the experimenter to integrate highly
specialized equipment in unique configurations, while often lacking expertise
in the relevant engineering fields.

Bonsai has a large user base in the systems neuroscience community.
Quantifications of this user base are provided in Figures 1d-1f.  In the last
year alone, we estimate more than 1,000 new users have started to incorporate
Bonsai into their experimental protocols across the world. The surprising rate
of adoption of Bonsai in non-programmer experimental labs highlights the need
for accessible programming tools that enable state-of-the-art technology but
also allow researchers to stay in control and change their experimental
paradigms.  Many open-source software tools are either inaccessible to
non-programmers, or too constrained to be of general use outside their narrow
domain of application. Bonsai has been successful because it has straddled this
gap to some extent.

The language has also helped to potentiate the growing wave of foundational
open hardware initiatives, such as the OpenEphys \citep{siegleEtAl17} and UCLA
Miniscope~\citep{caiEtAl16}, making it possible to quickly combine and
integrate these tools into new experiments~\citep{buccinoEtAl18}.
%
Bonsai has been adopted in large neuroscience undertakings like the
International Brain
Laboratory\footnote{\href{https://www.internationalbrainlab.com/}{https://www.internationalbrainlab.com/}}
and the Allen Institute for Brain
Science\footnote{\href{https://alleninstitute.org/what-we-do/brain-science/}{https://alleninstitute.org/what-we-do/brain-science/}}.
%
More recently, Bonsai has started to expand outside the domain of neuroscience
into biomedical research and biotechnology tool development, and even outside
academia into public outreach and education programs.

\subsubsection*{Other resources in the subject area}

The space of technologies serving experimental control and behaviour monitoring
is large, and is traditionally occupied either by domain-specific graphical
user interfaces for control and recording of specific devices and experiment
types (e.g. Open Ephys
GUI\footnote{\href{https://open-ephys.org/gui/}{https://open-ephys.org/gui/}},
Miniscope DAQ
Software\footnote{\href{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software}{https://github.com/Aharoni-Lab/Miniscope-DAQ-QT-Software}})
or by real-time control frameworks for specifying task logic using state
machine or similar formalisms (e.g. NIMH ML
(nih.gov)\footnote{\href{https://monkeylogic.nimh.nih.gov/}{https://monkeylogic.nimh.nih.gov/}},
pyControl\footnote{\href{https://pycontrol.readthedocs.io/en/latest/}{https://pycontrol.readthedocs.io/en/latest/}},
Autopilot\footnote{\href{https://docs.auto-pi-lot.com/en/latest/}{https://docs.auto-pi-lot.com/en/latest/}},
Sanworks\footnote{\href{https://sanworks.io/index.php}{https://sanworks.io/index.php}}).
These dedicated interfaces are typically very comfortable for experimenters
operating within the specific domain that the tool is designed for but tend to
become unwieldy when there is a need to introduce a new device or variation of
a task which is outside the scope of the framework. The alternative is to use a
more general programming language like Python or MATLAB, with the disadvantage
of the code being harder to understand, maintain, and change. Programming
languages like LabVIEW straddle the middle ground and provide a high-level and
flexible visual interface for composing data acquisition and control systems
themselves. Unlike Bonsai, however, the graphical elements of LabVIEW are
heterogeneous and very fine grained, creating the need for long and complicated
logical structures to implement even a simple experimental control system. By
providing an extremely simple, yet flexible, visual syntax, Bonsai opens the
opportunity even for complete non-programmers to design and successfully
customize relatively complex experiments from the ground up. It is mostly this
capability which has made Bonsai so attractive as a standard tool in
experimental neuroscience.