### 2021-11-16 application/workPlan.tex
Added

\documentclass[a4paper,11point]{article}
\usepackage[margin=2cm]{geometry}
\usepackage{graphicx}

\makeatletter
\setlength{\@fptop}{0pt}
\makeatother

\author{}
\date{}

\begin{document}

\begin{figure}[!t]
   \begin{center}
       \includegraphics[width=7in]{figures/workPlan_coarse.png}

       \caption{Gnatt chart for major project tasks. The label next to the task
       name refers to the section of \emph{Case for Support} document where the
       task is described}

     \label{fig:gnattChart}
   \end{center}
\end{figure}

\end{document}