### 2021-11-16 application/trackRecord.tex

The team responsible for the proposed project comprises two academic
institutes at University College London, the Gatsby Computational Neuroscience Unit (GCNU) and the Sainsbury
Wellcome Centre for Neural Circuits and Behavior (SWC), and a private
collaborator, NeuroGEARS Ltd.

The GCNU is a world-class research center for computational neuroscience and
machine learning, while the Sainsbury Wellcome Center is a top-notch experimental
neuroscience research center. They share a common building and collaborative
activities are at the heart of both institutes.
%
NeuroGEARS is a spinoff of the SWC and is currently contributing to a major
collaborative research project between the SWC and the GCNU. It has developed
Bonsai, a very unique ecosystem for experimental control and data acquisition,
it has offered courses around the globe instructing students on advanced
experimental control, and has used Bonsai in numerous research and private
experimental projects. Thus, NeuroGEARS has extensive experience on the
implementation of neuroscience experiments.

In summary, the partners in this project have outstanding expertise on their
focus areas (computational neuroscience and machine learning, experimental
neuroscience, implementation of experiments), as well as a rich collaborative
track record. The proposed project will merge this unique expertise to create a
resource that will be essential for the advanced of biological sciences.