### 2021-11-16 application/trackRecord.tex

The team responsible for the proposed project comprises two academic
institutes at University College London, the Gatsby Computational Neuroscience Unit (GCNU) and the Sainsbury
Wellcome Centre for Neural Circuits and Behaviour (SWC), and a private
collaborator, NeuroGEARS Ltd.

The GCNU is an authority in computational neuroscience and
machine learning, while the Sainsbury Wellcome Centre is a world-leading experimental
neuroscience research institute. They share a common building, which facilitates regular interaction and collaboration across various research groups, including the two named PIs in this project, and also collaborate in many joint activities.
%
NeuroGEARS is a spinoff of the SWC and is currently contributing to a major
collaborative research project between the SWC and the GCNU. It developed
Bonsai, a unique ecosystem for experimental control and data acquisition,
and offers courses around the globe instructing students on advanced
experimental control. NeuroGEARS has used Bonsai in numerous research and private experimental projects, and has extensive experience on the
implementation of neuroscience experiments.

In summary, the partners in this project have outstanding expertise on their focus areas (computational neuroscience and machine learning, experimental
neuroscience, implementation of experiments), as well as a rich collaborative
track record. The proposed project will merge this unique expertise to create a
resource that will be essential for the advanced of biological sciences.