### 2021-11-17 application/figures/extensions.tex

% \documentclass{standalone}
% %% Fonts, etc.

% %% Language and font encodings
% \usepackage[english]{babel}
% \usepackage[utf8x]{inputenc}
% \usepackage[T1]{fontenc}

%  \usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
%  \usepackage{xcolor}
% % \usepackage{soul} % \ul \st \hl

% %% Maths

% % \usepackage{amssymb,amsmath}
% % \input m.macros
% % \input m.symbols

% %% Graphics

% \usepackage{graphicx,tikz}
% \usetikzlibrary{positioning}
% \usetikzlibrary{fit}
% \usetikzlibrary{calc}

% %% Document
% \begin{document}

\begin{tikzpicture}[
    ML/.style={text width=10em, rounded corners,
      fill=green!50!black!30, draw=green!50!black, very thick},
    Ext/.style={text width=10em, rounded corners,
      fill=blue!50!black!30, draw=blue!50!black, very thick}
    ]

  \node[text width=10em, font={\bf}, align=center]
       (bonsai)
       {\includegraphics[width=7em]{bonsai-lettering}\\
         Bonsai core (C\#)};

 \coordinate [below = 5em of bonsai.south] (A);

  \node[below left=0 and 1em of A, Ext]
       (lang)
       {\textbf{Language interface:}\\
         C\# $\leftrightarrow$ Python\\
         $\qquad \leftrightarrow$ R\\
         $\qquad \leftrightarrow$ MATLAB
       };

  \node[below right=0 and 1em of A, Ext]
       (abstr)
       {\textbf{Analysis layers:}\\
         \texttt{Estimator}\\
         \texttt{Predictor}\\
         \texttt{Tranformer}};

  \node[above=1ex of A, text=blue!50!black, font={\bf}, inner sep=0]  (intr) {Interface layers};

  \node[fit=(lang) (abstr) (intr), line width=1mm, draw=blue!50!black!50, rounded
    corners, inner sep=1ex] (intr box) {};

 \coordinate [below = 12em of A] (B);

 \node[below=0 of B, ML]
       (signal)
       {\textbf{Signal analysis:}\\
         Spike detection\\
         LFP extraction\\
         Decoding};

  \node[below left=0 and 1em of signal.north west, ML]
       (video)
       {\textbf{Video analysis:}\\
         Motion sequencing\\
         Behaviour decoding\\
         Interpretable latents};

  \node[below right=0 and 1em of signal.north east, ML]
       (dimred)
       {\textbf{Population analysis:}\\
         Gaussian LDS\\
         Poisson LDS\\
         HMM\\
         GPFA\\
         svGPFA\\
         LFADS};

  \node[above=1ex of B, text=green!50!black, font={\bf}, inner sep=0] (ml)
       {Machine-intelligence algorithms};

  \node[fit=(video) (signal) (dimred) (ml), line width=1mm, draw=green!50!black!50, rounded
    corners, inner sep=1ex] (ml box) {};

  \draw[<->, line width=1mm, gray] (bonsai) -- (intr box);

  \draw[<->, line width=1mm, gray] (intr box) -- (ml box);

  %% \node[below = 15em of bonsai, ML]
  %% (compare)
  %%      {\textbf{Method comparison}};

\end{tikzpicture}

% \end{document}

% Local Variables:
% mode: latex
% eval: (set-fill-column 80)
% End: