### 2021-11-17 application/Je-S_sections.tex

List the main objectives of the proposal in order of priority [up to
  4000 chars]

The overall goal of this proposal is to extend, enhance, maintain and support Bonsai, a fully integrated software environment to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models.

Specifically, we will:

1. Develop, test and deploy a new software workflow to incorporate external packages into Bonsai's C# environment. \textcolor{red}{I believe this is already done. All existing bonsai community packages already depend on the existence of the extendable plugin architecture -- perhaps you mean "languages" rather than "packages"? i.e. MATLAB, Python, R, in that case I would rewrite as below}

1. Develop, test and deploy new foreign language interfaces to allow extending the Bonsai environment with new packages developed in languages other than C#, namely MATLAB, Python and R.

2. Define a standard API and set of operators appropriate for composing machine-learning functionality into a Bonsai workflow, specifically to enable:
2a. learning
2b. inference
2c. visualisation
2d. model-comparison and validation
2e. exchange of models and hyper-parameters

3. Incorporate a suite of machine-learning-driven analysis tools into the Bonsai environment using the standard API specified above, to implement:
3a. video-based behavioural analysis
3b. real-time interfacing of population-scale extracellular activity with laboratory control  
3c. evaluation of competing data analytic approaches within a single experimental framework

4. Enhance the quality and reproducibility of systems neuroscience through:
4a. fully open-source and public-facing development
4b. rapid release of, and community engagement around, new tools
4c. provision of standard implementations of analysis and control methods, with the ability to distribute fully automated and reproducible analysis pipelines.

5. Extend and support the Bonsai user community through code and forum engagement and development workshops.

[2829 chars remaining]

Describe the Bioinformatic and/or Biological Resource in simple terms in a way that could be publicised to a general audience [up to 4000 chars]

Understanding the brain is a major scientific challenge of our era. To succeed, scientists must be able to explain how animal behaviour relates to neural activity across different brain regions. This requires careful design and manipulation of behavioural experiments, where experimenters either record or manipulate neural activity while the animals (e.g. non-human primates, rodents, fish, insects) perform particular tasks.

Experiments in behavioural and brain science laboratories require software that integrates and controls hardware from multiple recording devices (video, electrodes for neural activity measurement, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets.

Scientists studying brain and behaviour dedicate the majority of their time designing experiments and analysing the data, with least amount spent on data acquisition itself, which may impact the quality of data. Moreover, hundreds of neuroscience research groups worldwide develop their own experimental and analytical tools, most using different programming languages, leading to  inefficiencies in data sharing and analysis, and impacting reproducibility (i.e. how easy it is for someone else to do the same experiment).

Here we propose to provide the neuroscience community with a software tool that will dramatically increase the efficiency of experimental control and data analysis. We will do so by developing a new set of functionalities to an existing software platform, Bonsai. Bonsai is a fully integrated software environment that emphasises performance, flexibility, and ease-of-use, allowing scientists with no previous programming experience to quickly develop their own high-performance data acquisition and experimental control systems.

Thus far, Bonsai has been adopted by hundreds of scientists worldwide to provide interactive experimental control in behavioural and brain sciences. In this proposal, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural datasets, and to create an open-access platform for software sharing. Bonsai’s enhanced functionality will enable new types of research, and speed up discovery and improve efficiency by (i) providing access to such tools to laboratories lacking expertise, (ii) reducing the need to reinvent the same tools in multiple labs and (iii) standardising the data processing streams, thus increasing reproducibility across laboratories. We believe this effort will enable and accelerate new discoveries in how the brain generates behaviour.

Describe the Bioinformatic and/or Biological Resource in a manner suitable for a specialist reader. This summary will be made publicly available if the proposal is funded. [up to 2000 characters]

To understand the brain, scientists aim to explain how animal behaviour relates to neural activity. This requires the design and precise control of behavioural experiments, wherein animals perform particular tasks while experimenters either record or manipulate neural activity in specific neural circuits. Such experiments require data acquisition software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex datasets. Progress is held back by the lack of standardised tools for design and implementation of experimental protocols, and the difficulty of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs. The fields of behavioural and brain sciences have consequently suffered from both inefficiency and poor reproducibility, due to disparate data acquisition and analysis solutions created independently across laboratories.

To address these challenges, we propose to extend, enhance, maintain and support \textbf{Bonsai}, a fully integrated software environment to enable cutting-edge reproducible systems neuroscience experiments using animal models, with a particular emphasis on machine-intelligence-enabled, real-time neuroinformatics methods. While Bonsai is already adopted by hundreds of scientists worldwide, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and offline analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages. Enhancing Bonsai's ecosystem will be a game-changer for behavioural and brain science experiments by enabling new types of research, increasing and diversifying user base, and dramatically improve efficiency and reproducibility of research.

Describe who will benefit from this Bioinformatic and/or Biological resource(s) being funded [up to 4000 characters]

The main beneficiaries of this proposal will be experimental labs worldwide who currently lack the highly technical expertise required to apply Machine Intelligence to their experiments. While more and more automated Machine Learning platforms are becoming generally available across many data science fields, they either focus on making available completely general-purpose tools which require deep expertise and training to construct and train new algorithms and models from scratch, or they encapsulate some pre-trained and domain-specific model which is useful only for very specific classes of problems (e.g. face recognition, pose estimation).

A middle ground where existing algorithms can be easily composed by non-experts in a flexible way to design entirely novel Machine Intelligence applications tailor-made for specific research questions is currently entirely absent. In fields like systems neuroscience, where progress is made essentially by exploring the vast unknown space of animal behaviour and wide-scale brain physiology, the need for new bespoke algorithms and analysis procedures is constant and critical. If successful, this proposal will entirely transform the landscape of Machine Intelligence applied to experimental lab research in biology, by empowering non-technical researchers to recombine and repurpose Machine Intelligence tools using a flexible visual programming environment and providing seamless integration with all the measurement, instrumentation and control packages already available in Bonsai.

In addition to the researchers themselves, the standard analysis pipelines will allow everyone who is interested in asking questions about the behaviour and physiology of animals to analyse and understand their data. As an example, increasingly in research institutions worldwide Bonsai is being used to monitor animal welfare by in-house technicians who are concerned about a broader range of environmental variables which might impact animal housing quality but which are beyond the ones which are available in standard vivarium solutions. They use Bonsai as it allows them to prototype and design entirely new monitoring systems on-site, without the need to engage an expensive software development process to answer simple questions. With access to the neuroinformatics packages developed for this proposal, staff at these institutions would be immediately empowered to address more sophisticated questions regarding interactions between housing conditions and animal physiology, which at the moment are entirely unthinkable, but would lead to instantly improving the welfare of hundreds of thousands of research animals in biological research facilities.