### 2021-11-17 application/caseForSupport.tex
Updates from Overleaf

\documentclass[a4paper,11pt]{article}
\usepackage[margin=2cm]{geometry}
%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
% \usepackage{microtype}      % fine pdf font control

\usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
\usepackage{xcolor} % -- loaded by tikz

% \usepackage[colorlinks=true]{hyperref}
\usepackage[colorlinks = true,
            linkcolor = blue!50!black!50,
            urlcolor  = blue!50!black!50,
            citecolor = blue!50!black!50,
            anchorcolor = blue!50!black!50]{hyperref}
\usepackage[super]{natbib}
\renewcommand{\citet}[1]{[\citenum{#1}]}
% \usepackage{apalike}
\usepackage{graphicx}
\usepackage{verbatim}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{calc}

\usepackage{float}
\usepackage{wrapfig}

\usepackage[font={footnotesize}]{caption}

\makeatletter
\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
            {-2.5ex\@plus -1ex \@minus -.25ex}%
            {1.25ex \@plus .25ex}%
            {\normalfont\normalsize\bfseries}}
\makeatother
\setcounter{secnumdepth}{4} % how many sectioning levels to assign numbers to
\setcounter{tocdepth}{4}    % how many sectioning levels to show in ToC
\usepackage[tiny]{titlesec} % format section titles
\def\headfmt{\color{blue!50!black!50}\bfseries}
\titleformat*{\section}{\headfmt}
\titlespacing*{\section}{0pt}{*1}{*0}[-4em]
\titleformat*{\subsection}{\headfmt}
\titlespacing*{\subsection}{0pt}{*0.5}{*0}
\titleformat*{\subsubsection}{\headfmt}
\titlespacing*{\subsubsection}{0pt}{*0}{*0}
\titleformat*{\paragraph}{\headfmt}
\titlespacing*{\paragraph}{0pt}{*0.5}{*0}

\usepackage{enumitem}
\setlist[description]{
    topsep=0pt,
    itemsep=0pt,
    partopsep=0pt,
    parsep=0pt,
    font=\headfmt,
    leftmargin=1em}

\renewcommand{\footnote}[1]{ [#1]}

\title{Case for Support}
\author{}

\begin{document}

% \maketitle

\tableofcontents

\newpage

\setcounter{page}{1}

\section*{\large Case for support}
\vskip 0.5ex
We propose to extend, enhance, maintain and support \textbf{Bonsai}, an integrated software environment to enable \textbf{cutting-edge reproducible
  laboratory systems neuroscience experiments in animal models}, with a
particular emphasis on \textbf{machine-intelligence-enabled, real-time
  neuroinformatics} methods.

\input{backgroundToTheResource}

\input{detailsOfResource}

\section{Community demand}
\label{sec:communityDemand}

\input{communityDemand}

\section{User engagement}
\label{sec:userEngagement}

\input{userEngagement}

% \section{Resource management}

\input resourceManagementandSustainability

% \section{Long term sustainability planning}

\section{Potential for economic and societal impact}

\begin{description}

    \item[Understanding the brain and behaviour] Systems neuroscientists aim to
        uncover how adaptive behaviour arises from computation in neural
        circuits in different brain regions. Extending the Bonsai ecosystem
        will facilitate new types of experiment that causally connect neural
        circuits with elemental behavioural processes. Long-term, this research
        will help identify missing links between circuit dysfunction and
        behavioural deficits in neurological or mental health disorders, and
        will support the development of new neural circuit biomarkers
        and therapies. The Bonsai ecosystem will also enable  experimental
        testing of emerging theories of brain function and learning, and thus
        contribute to the neuro-inspired machine learning algorithms that fuel
        intelligent technologies, increasingly important in our society.  The
        enhanced Bonsai platform will also support purely behavioural
        experiments, which track animals' posture, body parts or location in real-time, to investigate impacts of environmental variables
        or captivity on behaviour - crucial for
        animal welfare.

    \item[Improve reproducibility in behavioural and brain sciences] Lack of
        reproducible results slows down scientific progress and incurs personal
        and economic costs to scientists and funders.  In neuroscience,  many
        behavioural and neural animal studies have been difficult to
        reproduce, and experiments are rarely replicated across laboratories.
        By creating a resource that standardises and enables reproducible
        experimental design/control, as well as analysis of neural and
        behavioural data, the Bonsai ecosystem will allow multiple laboratories
        to implement and align experimental and analytical protocols as closely
        as possible.

    \item[Improve efficiency of research] Most neuroscience laboratories
        worldwide develop custom, unique solutions to their experimental
        control and analysis needs. Some laboratories also lack the technical
        know-how to either develop their own software for experimental hardware
        control or write software for state-of-the-art analysis of rich
        neural and behavioural data, increasingly reliant on latest machine
        learning algorithms. The proposed upgrading of the open-source and
        community-empowered Bonsai ecosystem with a toolbox of online and
        offline analytical tools, which can dovetail with closed-loop
        experimental requirements, will (i) allow the deployment of such tools
        into laboratories that lack the necessary expertise, (ii) reduce the need to
        reinvent the same tools in multiple labs (cost/time efficiency) and
        (iii) concomitantly streamline and standardise the experimental data
        processing streams in systems neuroscience.

\end{description}

\section{Track Record}

\input{trackRecord}

\bibliographystyle{unsrtveryabbrv}
\bibliography{journalsabbrvnodots,bonsai,epsrcSoftwareForResearchCommunities,monitoringBehavior,stateSpaceModels,machineLearning,gaussianProcesses,spikeSorting}

\end{document}