### 2021-11-17 application/trackRecord.tex

The proposed project team comprises two academic
institutes at University College London, the Gatsby Computational Neuroscience Unit (GCNU) and the Sainsbury
Wellcome Centre for Neural Circuits and Behaviour (SWC), and a private
collaborator, NeuroGEARS Ltd.

The GCNU is an authority in computational neuroscience and
machine learning, while the SWC is a world-leading experimental
neuroscience research institute. They share a common building, which facilitates regular interaction and collaboration across research groups, including the two named PIs in this project, and also collaborate in many joint activities.
%
NeuroGEARS is a consulting company started by the creators of Bonsai. It is currently contributing to a major
collaborative research project between the SWC and the GCNU. It continues to develop Bonsai,
and offers courses around the globe instructing students on advanced experimental control.
NeuroGEARS has used Bonsai in numerous research and private experimental projects,
and has extensive experience on the implementation of neuroscience experiments.

In summary, the partners in this project have outstanding expertise on their focus areas (computational neuroscience and ML, experimental
neuroscience, implementation of experiments), as well as a rich collaborative
track record. The proposed project will merge this unique expertise to create a
resource that will be essential for the advanced of biological sciences.