### 2021-11-17 application/bonsai.tex

\textbf{Bonsai}~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual programming language developed 
in response to these challenges.
% ~\citep{lopesEtAl15,lopesAndMonteiro21} is a free and open-source visual
% programming language that 
Its design emphasises performance, flexibility, and ease-of-use,
allowing scientists with no previous programming experience to quickly develop
their own high-performance data acquisition and experimental control systems.
Bonsai combines a high-level event algebra for data streams with an integrated
development environment (IDE) and an extensive library of plugins supporting
multiple hardware and software packages used by the neuroscience research
community (Figure~\ref{fig:bonsai}A-B).
%
A Bonsai graphical program consists of one or more source data streams (e.g.\ neural activity, video, audio, sensors, etc)
and several interconnected operators that transform input to output
datastreams (Figure~\ref{fig:bonsai}C).

% I think we've said something similar in the first para  - m.
%
% Standard software tools for data analysis (ImageJ, MATLAB, R, etc.) have
% transformed progress inf increasingly “data-rich” sciences. However,
% equivalent standardised software tools for data acquisition and control of
% animal behaviour in neuroscience experiments are still lacking. Life science experiments demand
% a combination of multiple instrumentation and control technologies, for both
% behavioural and physiological investigations. The growing complexity on both the
% amount of data that is collected, and the rich conditions under which behaviour
% must be explored, place an increasing burden on experimenters to integrate
% highly specialised equipment in unique configurations, while often lacking
% expertise in the relevant engineering fields.

Bonsai has been adopted in hundreds of laboratories worldwide and has the largest user base in the systems neuroscience community (Figure~\ref{fig:bonsai}D-F).  In the last year alone, more than 1,000 new users incorporated Bonsai into their experimental protocols. The rapid rate of adoption of Bonsai in non-programmer experimental
labs highlights the need for accessible design tools that enable
state-of-the-art technology but also allow researchers to stay in control and flexibly
change their experimental paradigms.  Many open-source software tools are
either inaccessible to non-programmers, or too constrained to be of general use
outside their narrow domain of application. Bonsai has been successful because
it bridges this gap.

Bonsai also supports the growing wave of foundational
open hardware initiatives, such as OpenEphys \citep{siegleEtAl17} and UCLA
Miniscope~\citep{caiEtAl16}, allowing these tools to be quickly combined and integrated into new experiments~\citep{buccinoEtAl18}.
%
Bonsai has been adopted in large neuroscience undertakings like the
International Brain
Laboratory\footnote{https://www.internationalbrainlab.com/}
and the Allen Institute for Brain
Science\footnote{https://alleninstitute.org/what-we-do/brain-science/}.
%
Key to the wide adoption is the reproducibility that Bonsai ensures.  Data acquisition and experimental control protocols can be 
replicated in any laboratory just by sharing a Bonsai configuration file. 
%
% LEAVE OUT FOR THIS GRANT:
% More recently, Bonsai has started to expand outside the domain of neuroscience
% into biomedical research and biotechnology tool development, and even outside
% academia into public outreach and education programs.