### 2021-11-17 application/resourceManagementandSustainability.tex

\section{Resource management}
%
As an open source project developed over almost 10 years, Bonsai 
has well-established management practices built around the user community.
%
The source code for the core system and the majority of extension packages
are hosted at \url{github.com/bonsai-rx}.
%
The documentation, including learning materials and all relevant community
links are publicly accessible at \url{bonsai-rx.org}.
%
A community user forum is a public Google Group with more than 700 active users (Figure~\ref{fig:bonsai}f) exchanging several
new questions and answers every day.
%
Package installation and updates are supported by standalone package
management software included within the core Bonsai distribution and hosted
in a shared centralised and curated repository.
%
This ensures that bugfixes and enhancements are easily available and, importantly, allows for reproducible deployment of user environments
to ensure working experiments are not disrupted by system changes.
Large-scale experimental projects such as the International Brain Laboratory
have leveraged this capability of Bonsai to standardise data acquisition and
control software across dozens of laboratories around the world.
%
Community technology previews are made available through the package manager,
letting early adopters opt-in to the latest features
before they are generally available. This allows the community
to report issues with new features on GitHub, thereby
distributing bug reporting and triage, and improving the quality of
stable releases.

The core compiler and IDE is extensively tested, and uses
a strong type system to greatly reduce bugs at compile-time. Best
practices for leveraging this infrastructure for building new packages are
in place.

The software engineers recruited to this project will work with our
collaborators at NeuroGEARS Ltd, to strengthen and maintain these
practices.  Once fully familiar with the Bonsai core architecture they
will participate in review of pull requests and other code contributions.

% To complement the management structure document, proposals may outline
% how the resource management and advisory structures will operate including
% outlining any review procedures.

% Additionally, opportunities for staff training and support to ensure their
% continuous development should be identified to safeguard the successful
% delivery of the resource.

\section{Long-term sustainability planning}

Bonsai is a freely available open-source resource, already widely adopted by
the systems neuroscience and behavioural science communities (see Fig~\ref{fig:bonsai}).  
%
Current contributors include other large open projects (OpenEphys, DeepLabCut),
and companies developing and supplying new hardware devices  (Neurophotometrics).
%
This community uptake, engagement, and commitment provides a foundation for
long-term sustainability.

NeuroGEARS Ltd is a keystone of this community and a key collaborator in the current proposal.
Founded by the original developer of Bonsai, it helps to support and expand the
user base well beyond the life of the current proposal. NeuroGEARS currently provides
quality control and curation for the core Bonsai code, and provides training and support
services for research institutions around the world who want to start using Bonsai.
NeuroGEARS also facilitates knowledge transfer between users and institutions adopting
Bonsai, by moderating the community user forums and collaborating with organisations
who want to build new innovative research projects on top of Bonsai, as in the
current proposal.

The specific contributions developed as part of this project will form new core
components and packages within the Bonsai ecosystem. In particular, the new language integration packages will support long-term sustainability, by dramatically
increasing Bonsai's interoperability with existing software development communities which are
already well established in basic science (MATLAB, Python, R).
%
Developing these new packages with input from the core advisory panel of Bonsai,
and early public distribution and user testing within the SWC and with planned workshops, will ensure
solid and broad user engagement from the beginning of the development cycle and will therefore
build momentum for their use, support and dissemination.
%
Close collaboration with the core maintainers of the Bonsai ecosystem at NeuroGEARS will
allow them to take over responsibility for maintenance and support of these packages beyond the
end of the project, following the same principles of open-source and free sharing of knowledge
that characterises the entire Bonsai ecosystem.