### 2021-11-28 application/dataManagementPlan_long.tex
Added

\documentclass[a4paper,11pt]{article}
\usepackage[margin=2cm]{geometry}
%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
% \usepackage{microtype}      % fine pdf font control

\usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
\usepackage{xcolor} % -- loaded by tikz

% \usepackage[colorlinks=true]{hyperref}
\usepackage[colorlinks = true,
            linkcolor = blue!50!black!50,
            urlcolor  = blue!50!black!50,
            citecolor = blue!50!black!50,
            anchorcolor = blue!50!black!50]{hyperref}
\usepackage[super]{natbib}
\renewcommand{\citet}[1]{[\citenum{#1}]}
% \usepackage{apalike}
\usepackage{graphicx}
\usepackage{verbatim}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{calc}

\usepackage{float}
\usepackage{wrapfig}

\usepackage[font={footnotesize}]{caption}

\makeatletter
\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
            {-2.5ex\@plus -1ex \@minus -.25ex}%
            {1.25ex \@plus .25ex}%
            {\normalfont\normalsize\bfseries}}
\makeatother
\setcounter{secnumdepth}{4} % how many sectioning levels to assign numbers to
\setcounter{tocdepth}{4}    % how many sectioning levels to show in ToC
\usepackage[tiny]{titlesec} % format section titles
\def\headfmt{\color{blue!50!black!50}\bfseries}
\titleformat*{\section}{\headfmt}
\titlespacing*{\section}{0pt}{*1}{*0}[-4em]
\titleformat*{\subsection}{\headfmt}
\titlespacing*{\subsection}{0pt}{*0.5}{*0}
\titleformat*{\subsubsection}{\headfmt}
\titlespacing*{\subsubsection}{0pt}{*0}{*0}
\titleformat*{\paragraph}{\headfmt}
\titlespacing*{\paragraph}{0pt}{*0.5}{*0}

\usepackage{enumitem}
\setlist[description]{
    topsep=0pt,
    itemsep=0pt,
    partopsep=0pt,
    parsep=0pt,
    font=\headfmt,
    leftmargin=1em}

\renewcommand{\footnote}[1]{ [#1]}

\title{Data Management Plan}
\author{}
\date{}

\begin{document}

\maketitle

\setcounter{section}{-1}

\section{Proposal name}

\textcolor{orange}{\textit{Exactly as in the proposal that the DMP accompanies}}

Machine intelligence for (neuroscience) experimental control

\section{Description of the data}
\subsection{Type of study}

\textcolor{orange}{\textit{Up to three lines of text that summarise the type of study (or studies) for which the data are being collected}}

Enhancement of an existing software resource, the Bonsai software
ecosystem for data acquisition and experimental control.

\subsection{Types of data}

\textcolor{orange}{\textit{Types of research data to be managed: quantitative or
qualitative, sequencing data, images, models, software, scripts, protocols, and
procedures}}

We propose to incorporate into the Bonsai software ecosystem state-of-the-art
machine-learning models, as well as capabilities to enable the communication
between Bonsai, written in C\#, and machine-learning applications written in
Python, R and Matlab.

\subsection{Origin of the data}

\textcolor{orange}{\textit{Are you creating new primary data (i.e. collecting or
generating data)? Or are you re-using already available sources, including to
create new data (e.g. a new dataset created from transformation or integration
of existing data)? If data is being re-used, is it publicly accessible?}}

We will not create primary data for the proposed project. However, as outlined
in \emph{Section 2.5 Testing with neuroscience data} of the \emph{Case for
Support} document, the large community of Bonsai users at the Sainsbuy Wellcome
Centre (SWC) will test the new functionality added to Bonsai prior to its
distribution. For these tests, this community will use their own datasets, its
including high-density Neuropixels physiological recordings, with optogentic
manipulations and high-resolution video monitoring of animal behaviour. These
datasets will not be made publicly available for the purposes of the proposed
project. However, we are building infrastructure to make them available in the
near future.

The proposed enhancements to the Bonsai ecosystem will impact a wide range of
existing and to-be-created datasets collected by open-source initiatives
already integrated into the Bonsai ecosystem. For example, datasets of feeding
behaviour in flies (collected with FlyPAD
sensors\footnote{\url{https://flypad.rocks/}}), electrophysiology datasets
(collected with products supported by
OpenEphys\footnote{\url{https://open-ephys.org/}}), datasets of vivo imaging of
brain activity (collected with the UCLA
Miniscope\footnote{\url{http://miniscope.org/}} or with NeuroPhotometrics
products\footnote{\url{https://neurophotometrics.com/}}), datasets using rich
2D or 3D visual environments (created with
BonVision\footnote{\url{https://bonvision.github.io/}}), high-density
electrophysiology datasets (recorded with NeuroPixels
probes\footnote{\url{https://www.neuropixels.org/}}), and animal behavioral
datasets (with body parts tracked with
DeepLabCut\footnote{\url{http://www.mackenziemathislab.org/deeplabcut}}).

\subsection{Format and scale of the data}

\textcolor{orange}{\textit{File formats, software used, number of records, databases, sweeps, repetitions... (in terms that are meaningful in your field of research). An indication of the size of data to be stored and shared.}}

The source code for the core Bonsai ecosystem and the majority of extension
packages are hosted at \url{github.com/bonsai-rx}.

\section{Data management, documentation, and curation}

\textcolor{orange}{\textit{Keep this section concise and accessible to readers who are not data management experts.  Focus on principles, systems, and major standards. Focus on the main kind(s) of study data.  Give brief examples and avoid long lists.}}

% documentation
The documentation, including learning materials and all relevant community
links are publicly accessible at \url{bonsai-rx.org}. A forum where Bonsai
users daily exchange questions and answers can be accessed at
\url{https://groups.google.com/g/bonsai-users}. Currently more than 700 users
contribute to this forum (Figure~1f, \emph{Case for Support}).

% management
Package installation and updates are supported by standalone package management
software included within the core Bonsai distribution and hosted in a shared
centralised and curated repository. This ensures that bugfixes and enhancements
are easily available and, importantly, allows for reproducible deployment of
user environments to ensure working experiments are not disrupted by system
changes.
%
Large-scale experimental projects such as the International Brain Laboratory
have leveraged this capability of Bonsai to standardise data acquisition and
control software across dozens of laboratories around the world.
%
Community technology previews are made available through the package manager,
letting early adopters opt-in to the latest features before they are generally
available. This allows the community to report issues with new features on
GitHub, thereby distributing bug reporting and triage, and improving the
quality of stable releases.

The core compiler and IDE is extensively tested, and uses a strong type system
to greatly reduce bugs at compile-time. Best practices for leveraging this
infrastructure for building new packages are in place. The software engineers
recruited to this project will work with our collaborators at NeuroGEARS Ltd,
to strengthen and maintain these practices. Once fully familiar with the Bonsai
core architecture they will participate in review of pull requests and other
code contributions.

\subsection{Managing, storing and curating data}

\textcolor{orange}{\textit{Briefly describe how data will be stored, backed-up, managed and curated in the short to medium term. Specify any community agreed or other formal data standards used (with URL references).}}

Bonsai is a freely available open-source resource, already widely adopted by
the systems neuroscience and behavioural science communities (see Fig 1f).
Current contributors include other large open projects (OpenEphys, DeepLabCut),
and companies developing and supplying new hardware devices
(Neurophotometrics). This community uptake, engagement, and commitment provides
a foundation for long-term sustainability.

NeuroGEARS Ltd.\ is a keystone of this community and a key collaborator in the
current proposal. Founded by the original developer of Bonsai, it helps to
support and expand the user base well beyond the life of the current proposal.
NeuroGEARS currently provides quality control and curation for the core Bonsai
code, and provides training and support services for research institutions
around the world who want to start using Bonsai. NeuroGEARS also facilitates
knowledge transfer between users and institutions adopting Bonsai, by
moderating the community user forums and collaborating with organisations who
want to build new innovative research projects on top of Bonsai, as in the
current proposal.

The specific contributions developed as part of this project will form new core
components and packages within the Bonsai ecosystem. In particular, the new
language integration packages will support long-term sustainability, by
dramatically increasing Bonsai's interoperability with existing software
development communities which are already well established in basic science
(MATLAB, Python, R). Developing these new packages with input from the core
advisory panel of Bonsai, and early public distribution and user testing within
the SWC and with planned workshops, will ensure solid and broad user engagement
from the beginning of the development cycle and will therefore build momentum
for their use, support and dissemination. Close collaboration with the core
maintainers of the Bonsai ecosystem at NeuroGEARS will allow them to take over
responsibility for maintenance and support of these packages beyond the end of
the project, following the same principles of open-source and free sharing of
knowledge that characterises the entire Bonsai ecosystem.

\subsection{Metadata standards and data documentation}

\textcolor{orange}{\textit{What metadata is produced about the data generated from the research? For example descriptions of data that enable research data to be used by others outside of your own team.  This may include documenting the methods used to generate the data, analytical and procedural information, capturing instrument metadata alongside data, documenting provenance of data and their coding, detailed descriptions for variables, records, etc.}}

The fact that Bonsai is distributed as an open source project
(\url{https://bonsai-rx.org/}), that is has extensive online documentation
(\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
(\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
enables the use of the Bonsai ecosystem by the general public.

\end{document}