### 2021-11-29 application/dataManagementPlan_long_joaco.tex
Several modifications wrt dataManagementPlan.tex

\documentclass[a4paper,11pt]{article}
\usepackage[margin=2cm]{geometry}
%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
% \usepackage{microtype}      % fine pdf font control

\usepackage[scaled=0.92]{helvet} \renewcommand{\familydefault}{\sfdefault}
\usepackage{xcolor} % -- loaded by tikz

% \usepackage[colorlinks=true]{hyperref}
\usepackage[colorlinks = true,
            linkcolor = blue!50!black!50,
            urlcolor  = blue!50!black!50,
            citecolor = blue!50!black!50,
            anchorcolor = blue!50!black!50]{hyperref}
\usepackage[super]{natbib}
\renewcommand{\citet}[1]{[\citenum{#1}]}
% \usepackage{apalike}
\usepackage{graphicx}
\usepackage{verbatim}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{calc}

\usepackage{float}
\usepackage{wrapfig}

\usepackage[font={footnotesize}]{caption}

\makeatletter
\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
            {-2.5ex\@plus -1ex \@minus -.25ex}%
            {1.25ex \@plus .25ex}%
            {\normalfont\normalsize\bfseries}}
\makeatother
\setcounter{secnumdepth}{4} % how many sectioning levels to assign numbers to
\setcounter{tocdepth}{4}    % how many sectioning levels to show in ToC
\usepackage[tiny]{titlesec} % format section titles
\def\headfmt{\color{blue!50!black!50}\bfseries}
\titleformat*{\section}{\headfmt}
\titlespacing*{\section}{0pt}{*1}{*0}[-4em]
\titleformat*{\subsection}{\headfmt}
\titlespacing*{\subsection}{0pt}{*0.5}{*0}
\titleformat*{\subsubsection}{\headfmt}
\titlespacing*{\subsubsection}{0pt}{*0}{*0}
\titleformat*{\paragraph}{\headfmt}
\titlespacing*{\paragraph}{0pt}{*0.5}{*0}

\usepackage{enumitem}
\setlist[description]{
    topsep=0pt,
    itemsep=0pt,
    partopsep=0pt,
    parsep=0pt,
    font=\headfmt,
    leftmargin=1em}

\renewcommand{\footnote}[1]{ [#1]}

\title{Data Management Plan}
\author{}
\date{}

\begin{document}

\maketitle

% \textcolor{orange}{\textit{The following template is provided to applicants to assist in the development of a Data Management Plan (DMP) to accompany a research proposal. The notes (in italics) provide further context and guidance for its completion and should be deleted prior to submission. Applicants should consider submission of a DMP which is proportional to the quantity and nature of the data. Where substantial data is generated from the research, it is expected that the DMP will be longer and more detailed than studies generating small amounts of data, where DMPs may be significantly shorter. In either case, the DMP should be a maximum of 3 pages in length.}}

\setcounter{section}{-1}

\section{Proposal name}

% \textcolor{orange}{\textit{Exactly as in the proposal that the DMP accompanies}}

Machine intelligence for (neuroscience) experimental control

\section{Description of the data}
\subsection{Type of study}

% \textcolor{orange}{\textit{Up to three lines of text that summarise the type of study (or studies) for which the data are being collected}}

The project will generate software modules appropriate for use in behavioural neuroscience experiments in animal models, and other types of study that involve a combination of reactive experimental (particularly behavioural) control and streamed data acquisition.

% Enhancement of an existing software resource, the Bonsai software
% ecosystem for data acquisition and experimental control.

\subsection{Types of data}

% \textcolor{orange}{\textit{Types of research data to be managed: quantitative or
% qualitative, sequencing data, images, models, software, scripts, protocols, and
% procedures}}

This is a software project.  We will produce and manage new modules to be
integrated into the Bonsai software ecosystem.  These will include extensions
to incorporate non-native code packages into Bonsai (Section 2.1, Case for
Support), a general abstraction framework for data processing (Section 2.2,
Case for Support), machine learning functionality added to Bonsai (Section 2.3,
Case for Support) and documentation.

% We propose to incorporate into the Bonsai software ecosystem state-of-the-art
% machine-learning models, as well as capabilities to enable the communication
% between Bonsai, written in C\#, and machine-learning applications written in
% Python, R and Matlab.

\subsection{Origin of the data}

% \textcolor{orange}{\textit{Are you creating new primary data (i.e. collecting or
% generating data)? Or are you re-using already available sources, including to
% create new data (e.g. a new dataset created from transformation or integration
% of existing data)? If data is being re-used, is it publicly accessible?}}

This is a software project.  New modules will be created and tested by software research engineers at UCL in consultation with staff at NeuroGEARS Ltd (a project collaborator).  These will be integrated into the existing Bonsai software ecosystem.

We will access experimental data during the testing phase, but will not seek to manage these data within the project.  Such data will have been collected in the course of other projects and will be managed in accordance with the original protocols.

% We will not create primary data for the proposed project. However, as outlined
% in \emph{Section 2.5 Testing with neuroscience data} of the \emph{Case for
% Support} document, the large community of Bonsai users at the Sainsbuy Wellcome
% Centre (SWC) will test the new functionality added to Bonsai prior to its
% distribution. For these tests, this community will use their own datasets, its
% including high-density Neuropixels physiological recordings, with optogentic
% manipulations and high-resolution video monitoring of animal behaviour. These
% datasets will not be made publicly available for the purposes of the proposed
% project. However, we are building infrastructure to make them available in the
% near future.

% The proposed enhancements to the Bonsai ecosystem will impact a wide range of
% existing and to-be-created datasets collected by open-source initiatives
% already integrated into the Bonsai ecosystem. For example, datasets of feeding
% behaviour in flies (collected with FlyPAD
% sensors\footnote{\url{https://flypad.rocks/}}), high-resolution zebrafish
% tracking datasets (collected with
% BonZeb\footnote{\url{https://github.com/ncguilbeault/BonZeb/}}),
% electrophysiology datasets (collected with products supported by
% OpenEphys\footnote{\url{https://open-ephys.org/}} or with NeuroPixels
% probes\footnote{\url{https://www.neuropixels.org/}})), datasets of in-vivo
% imaging of brain activity (collected with the UCLA
% Miniscope\footnote{\url{http://miniscope.org/}} or with NeuroPhotometrics
% products\footnote{\url{https://neurophotometrics.com/}}), datasets using rich
% 2D or 3D visual environments (created with
% BonVision\footnote{\url{https://bonvision.github.io/}}), and animal behavioral
% datasets (with body parts tracked with
% DeepLabCut\footnote{\url{http://www.mackenziemathislab.org/deeplabcut}}).

\subsection{Format and scale of the data}

% \textcolor{orange}{\textit{File formats, software used, number of records, databases, sweeps, repetitions... (in terms that are meaningful in your field of research). An indication of the size of data to be stored and shared.}}

Software modules will comprise source files in a number of programming languages: C\#, Python, MATLAB and R.  These will be accompanied by extensive documentation (discussed below).

NEED: Approximate lines of code??? (order of magnitude is fine)

NEED: testing files, makefiles?, other stuff,

% The source code for the core Bonsai ecosystem and the majority of extension
% packages are hosted at \url{github.com/bonsai-rx}.

\section{Data management, documentation, and curation}

\textcolor{orange}{\textit{Keep this section concise and accessible to readers who are not data management experts.  Focus on principles, systems, and major standards. Focus on the main kind(s) of study data.  Give brief examples and avoid long lists.}}

\subsection{Managing, storing and curating data}
% \textcolor{orange}{\textit{Briefly describe how data will be stored, backed-up, managed and curated in the short to medium term. Specify any community agreed or other formal data standards used (with URL references).}}

% code
The source code for the core Bonsai ecosystem and the majority of extension
packages are hosted at \url{github.com/bonsai-rx} (MIT license with no
copyright and licencing restrictions).

% management
Package installation and updates are supported by standalone package management
software included within the core Bonsai distribution and hosted in a shared
centralised and curated repository. This ensures that bugfixes and enhancements
are easily available and, importantly, allows for reproducible deployment of
user environments to ensure working experiments are not disrupted by system
changes.
%
Large-scale experimental projects such as the International Brain Laboratory
have leveraged this capability of Bonsai to standardise data acquisition and
control software across dozens of laboratories around the world.
%
Community technology previews are made available through the package manager,
letting early adopters opt-in to the latest features before they are generally
available. This allows the community to report issues with new features on
GitHub, thereby distributing bug reporting and triage, and improving the
quality of stable releases.

The core compiler and IDE is extensively tested, and uses a strong type system
to greatly reduce bugs at compile-time. Best practices for leveraging this
infrastructure for building new packages are in place. The software engineers
recruited to this project will work with our collaborators at NeuroGEARS Ltd,
to strengthen and maintain these practices. Once fully familiar with the Bonsai
core architecture they will participate in review of pull requests and other
code contributions.

\subsubsection*{Team responsibilities}

\begin{description}

    \item[Goncalo Lopes] Bonsai related software development,
        software releases, organization of Bonsai workshops
        and co-supervision of research software engineer (RSE).

    \item[Joaquin Rapela] machine learning (ML) related software development,
        interaction with ML functionality methods developers, validating of ML
        functionality with behavioral and neural recordings and co-supervision
        of RSE.

    \item[Maneesh Sahani] advise on ML.

    \item[Tom Mrsic-Flogel] advise on neuroscience topics and communication
        with the steering committee.

\end{description}

\subsection*{Resourcing and responsibility}

\begin{description}

    \item[software development model] we will use the Kanban software
        development model, which we both Dr.~Lopes and Dr.~Rapela are already
        using in a collaborative project at the SWC/GCNU.

    \item[weekly meetings] between Goncalo Lopes, Joaquin Rapela and the RSE
        will be an essential component of the software development process.
        These will ensure that every member of the team is informed about both
        the Bonsai integration and the machine learning developments and will
        make the software development team more reliable upon the absence of
        any of its members.

    \item[code review and pair-programming] sessions will be held on a monthly basis among Goncalo Lopes, Joaquin Rapela and the recruited software engineer to deliver readable code sharing a standard programming style.

\end{description}

\subsubsection*{Major risks}

\begin{description}

    \item[language integration] we proposed to use existing software libraries
        to enable the communication between Bonsai, written in C\#, and other
        machine learning algorithms, written in Python (Python.NET), R (R.Net)
        and Matlab (native Matlab .Net interface). We have not yet tested any
        of these existing libraries, and we may found that they don't function
        reliably enough for our purposes. If this were the case, we proposed to
        build our own interface protocol using the ZeroMQ messaging library.
        This risk will only delay the progress of the proposed project, but
        will not hinder the deployment of its proposed outputs.

    \item[disfunctional machine learning functionality] in Section 2.3 of the
        Case for Support document (Figure 2, green box) we proposed to
        integrate existing machine learning algorithms into the Bonsai
        ecsosystem. It could be the case that some of these algorithms do not
        work as reliably as needed for our resource. I such a case we will use
        another algorithm with similar functionality.  This risk should not be
        a major problem, since most of the machine learning functionality we
        proposed to incorporate into Bonsai has been developed by us, or we are
        experienced using it.

    \item[lack of interest from Bonsai users] We will minimize the probability
        of this problem by advertising broadly the new functionality upon
        release.  However, we don't think this would be a real problem, since
        Bonsai users have reapeatly asked the addition of machine learning
        functionality to Bonsai (Section 3 Community demand, Case for Support),
        and because the machine learning functionality we proposed to add to
        Bonsai are popular among experimental-control neurocientists.

    \item[lack of interest from machine-learning methods developers] To
        minimize the probability that this risk turns into a large problem we
        will take three actions. First, we will provide machine-learning
        methods developers and easy-to-use API to interface their methods into
        the Bonsai ecosystem. Second, we will show machine-learning methods
        developers the large user exposure that Bonsai provides to integrated
        ML functionality. Third, we will organize Bonsai workshops for methods
        developers showcasing machine learning functionality integrated into
        the Bonsai ecosystem and instructing them how to integrate new methods
        into Bonsai.

    \end{description}

% Bonsai is a freely available open-source resource, already widely adopted by the systems neuroscience and behavioural science communities (see Fig 1f).  Current contributors include other large open projects (OpenEphys, DeepLabCut), and companies developing and supplying new hardware devices (Neurophotometrics). This community uptake, engagement, and commitment provides a foundation for long-term sustainability.

% NeuroGEARS Ltd.\ is a keystone of this community and a key collaborator in the current proposal. Founded by the original developer of Bonsai, it helps to support and expand the user base well beyond the life of the current proposal.  NeuroGEARS currently provides quality control and curation for the core Bonsai code, and provides training and support services for research institutions around the world who want to start using Bonsai. NeuroGEARS also facilitates knowledge transfer between users and institutions adopting Bonsai, by moderating the community user forums and collaborating with organisations who want to build new innovative research projects on top of Bonsai, as in the current proposal.

% The specific contributions developed as part of this project will form new core components and packages within the Bonsai ecosystem. In particular, the new language integration packages will support long-term sustainability, by dramatically increasing Bonsai's interoperability with existing software development communities which are already well established in basic science (MATLAB, Python, R). Developing these new packages with input from the core advisory panel of Bonsai, and early public distribution and user testing within the SWC and with planned workshops, will ensure solid and broad user engagement from the beginning of the development cycle and will therefore build momentum for their use, support and dissemination. Close collaboration with the core maintainers of the Bonsai ecosystem at NeuroGEARS will allow them to take over responsibility for maintenance and support of these packages beyond the end of the project, following the same principles of open-source and free sharing of knowledge that characterises the entire Bonsai ecosystem.

\subsection{Metadata standards and data documentation}

% \textcolor{orange}{\textit{What metadata is produced about the data generated from the research? For example descriptions of data that enable research data to be used by others outside of your own team.  This may include documenting the methods used to generate the data, analytical and procedural information, capturing instrument metadata alongside data, documenting provenance of data and their coding, detailed descriptions for variables, records, etc.}}

% The fact that Bonsai is distributed as an open source project
% (\url{https://bonsai-rx.org/}), that is has extensive online documentation
% (\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
% (\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
% many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
% enables the use of the Bonsai ecosystem by the general public.

% documentation
The documentation, including learning materials and all relevant community
links are publicly accessible at \url{bonsai-rx.org}. A forum where Bonsai
users daily exchange questions and answers can be accessed at the public
repository \url{https://groups.google.com/g/bonsai-users}. Currently more than
700 users contribute to this forum (Figure~1f, \emph{Case for Support}).

Classes, methods and functions will be documented in the code, and automatic
documentation tools (e.g., Sphinx for Python code, or Sphinx for .Net for C\#)
will be used to automatically generate API documentation. API documentation
will be complemented with class and interaction diagrams, documenting the
structure and dynamics of the application objects.

\subsection{Data preservation strategy and standards}

\textcolor{orange}{\textit{Plans and place for long-term storage, preservation,
and planned retention period for the research data. Formal preservation
standards, if any. Indicate which data may not be retained (if any). Also,
briefly describe what will be stored long-term, raw data, processed data?}}

For persistent and long term preservation we propose to use
Zenodo\footnote{\url{https://zenodo.org/}}.

\section{Data sharing and access}
\subsection{Where will data be shared?}
% Identify any data repository (-ies) that are, or will be, entrusted with storing, curating and/or sharing data from your study, where they exist for particular disciplinary domains or data types.

Software modules and documentation will be available on Github, with links from
the Bonsai project site (https://bonsai-rx.org) and available through the
Bonsai system's package management interface.

NEED: github org and status?

\subsection{When will data be available?}
% It is expected that timely release would generally be no later than the release through publication of the main findings and should be in line with established best practice in the field. Where best practices do not exist release within three years of generation of the dataset is suggested as a guide.

% Release management: repositories will be public throughout, but marked as developmental.  Alpha-level code will be available but not publicised until after extensive testing.  Will be integrated into package management and advertised once it achieves release quality.

Method to accept a project output: software will be automatically tested with a
comprehensive set of unit test and integration tests, implemented as  GitHub
Actions. In addition, prior to distribution to the general public, Bonsai users
inside the SWC will evaluate the functionality of the software with their own
state-of-the-art datasets. Project outputs will be accepted when all tests run
successfully and no major functional problems are reported by SWC Bonsai users.

Release management: we will create a new release after each of the machine
learning functionalities (described in Section~2.3 of the Case for Support) is
integrated into the Bonsai ecosystem following the above method to accept a
project output. Credentials for release will be granted initially by Dr.
Gonçalo Lopes, Head Engineer at Neurogears and the current curator of the
Bonsai official package feed, but might be expanded to other team members after
training.  Acceptance criteria for release will be a group decision between
Dr.~Lopes, Dr.~Rapela, Prof.~Sahani and Prof.~Mrsic-Flogel.

\subsection{How will data be made findable and accessible?}

Package management
Bonsai website
Community fora
Workshops
Paper?

% Indicate how potential new users (outside of your organisation) can find out about your data and identify whether it could be suitable for their research purposes, also consider how it will be licenced. The approach should encompass data in all its forms as well as software. This section could also outline access procedures, how your data will be able to be cited and tracked.

\subsection{How will data be made reusable?}

This is an open source software project.  Code will be distributed under ??? license.

% What steps will you take to ensure that your data could be re-analysed by other researchers? (e.g. availability of meta data, worked examples, vignettes). How will metadata be made available?

\subsection{Restrictions or delays to sharing, with planned actions to limit such restrictions} 
Restriction to data sharing may be due to subject confidentiality or to ensure you can gain IP protection. Applicants should ensure they have obtained necessary clearances from relevant collaborators with regards to the content of the proposal including the data sharing plan in line with the BBSRC Research Grants Guide. Where possible please indicate when data will be made available following a period of restriction.

All distributed software will be developed as part of the project, based on published and unrestricted algorithms.  Thus it is unencumbered by IP restrictions.  Will be available as above.

4. Data security (where relevant)
4.1	Formal information/data security standards
% Identify formal information standards with which your study is or will be compliant. An example is ISO 27001. If your organisation is ISO compliant, please state the registration number.

Not applicable.

4.2	Main risks to data security
% All personal data has an element of risk. Summarise the main risks to the confidentiality and security of information related to human participants or users, the level of risk and how these risks will be managed. Cover the main processes or facilities for storage and processing of personal data, data access, with controls put in place and any auditing of user compliance with consent and security conditions. It is not sufficient to write ‘not applicable’ under this heading.

This is a freely-available open source software package, which does not contain participants or personal data.   User access to the repository is anonymous and user personal information will not be collected.

\end{document}