### 2022-03-08 reviews/responseReviewers.tex
Added first comment from Maneesh

\documentclass[11pt]{letter}

\usepackage[hypertexnames=false,colorlinks=true,breaklinks]{hyperref}
\usepackage[a4paper, margin=2cm]{geometry}

\begin{document}

\begin{letter}{
}

Thank you for the reviewers' feedback and the opportunity to respond to
comments.

Response to reviewer 175134370:

Regarding the specific comment on reaching new user communities: the proposed
expansion of the Bonsai ecosystem, and the hackathons, workshops etc. mentioned
in the proposal, will be disseminated via numerous methods and fora. Many of
these methods have been successfully used to promote Bonsai thus far, including
professional networks, social media, and engagement with open-source hardware
projects such as Open-Ephys and Neurophotometrics, both of which have actively
promoted Bonsai for use with their systems. As part of the proposed extension,
we will demonstrate Bonsai and its new machine learning functionality at both
experimental and computational neuroscience venues (e.g., BNA's Festival of
Neuroscience, SFN's annual meeting, FENS meetings, Neuromatch Academy,
NeuroIPS, CoSyNe), to engage experimental neuroscience users, as well as new
neural data analysis method developers.

From SWC and GCNU, we will put together a comprehensive communications plan to
engage new communities with Bonsai. We will use a range of communications
channels including owned (SWC and GCNU websites, newsletters), shared (organic
social media) and paid (targeted/boosted social media and paid search)
channels. This could include publishing and publicising longer-form content on
how Bonsai is currently used by the global neuroscience community (e.g. at the
Allen Institute, Janelia Research Campus, etc.) and how the proposed expansion
will enhance this.

We will also empower current users to act as advocates for Bonsai, thereby
maximising reach through existing networks. Group Leaders will promote Bonsai
at conferences during research talks, and we will make targeted approaches to
Group Leaders not currently using the tools who may both benefit from and be
valuable promoters of Bonsai. This will include laboratories both within
neuroscience and from other fields— potentially leveraging user contacts in the
biological and drug screening groups already utilising Bonsai.

\pagebreak

Response to reviewer 155296694:

We share the concerns of the reviewer regarding people using machine learning
methods without understanding them. Some current data-analysis software
packages allow users to fit complex statistical models by just clicking a few
buttons in a graphical user interface. And sometimes, this ease of use leads
users to apply these statistical models without sufficient understanding. We
want to avoid this problem in Bonsai.

We will invest substantial resources to educate Bonsai users about the
functionality of the machine learning methods added to Bonsai. In addition, we
will provide extensive written and video documentation explaining the
integrated methods. We are currently developing a technical reference  website
for Bonsai which is open-source, integrated into the language editor, and open
for community contributions. We aim to make extensive use of this website to
explain relevant aspects of machine learning methods to the Bonsai community.
Furthermore, in collaboration with method authors, we will deliver talks at
workshops explaining the added methods.
%
The Gatsby Computational Neuroscience Unit is a centre of excellence in
statistical neuroscience and machine learning. Its contributions will be
essential to building simple and accurate statistical documentation related to
the interfaced methods.

Besides describing clearly in the documentation, and in the graphical user
interface, if suitable, the assumptions of every statistical method, in some
cases we may be able to add automatic tests checking if assumptions of a
statistical method (e.g., gaussianity, or homoscedasticity) are satisfied by the
data on which the method is applied.

In addition, for every machine-learning method that we will integrate into
Bonsai, we will provide recommendations on how to report analysis results.
This documentation will be targeted to methods' users, as well as journal and
grant reviewers. It will inform these reviewers what they could expect from
statistical analysis performed with methods integrated into the Bonsai
ecosystem.

Critical to helping Bonsai users understand the function of the integrated
machine learning methods, and interpret their estimated models' parameters, is
to provide extensive visualisations of these parameters. We will provide such
visualisations by taking advantage of Bonsai's built-in data visualisation
functionality.

Our vision is that Bonsai will not only allow to run advanced machine-learning
methods, but it will also serve as an educational tool, that will provide users
deep understanding of machine learning functionality for experimental control
and neural data analysis.

Regarding the comment about the lower uniqueness of our proposal by the fact that
the machine learning functionality that we proposed to add to Bonsai already
exists as free software package, we would like to clarify that it is not the
aim of the proposed resource to develop new machine learning functionality. We
intend to build a resource that can easily integrate into the experimental
control loop existing reliable machine learning functionality developed by the
machine learning community. In this way, the proposed resource should provide to
experimental neuroscientists up-to-date machine learning functionality, even
after the finalisation of the proposed development, in a sustainable manner. In
addition, the resource should provide methods developers rich datasets on which
to assess the performance of their contributed methods.  We believe that these
contributions are quite unique.

\pagebreak

A response to reviewer 206073742 is not required.

\end{letter}

\end{document}